SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-09-25 maintex.tex
updated discussion

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/prof_comp.pdf}
\caption[]{The mean radial profile of $\Delta$log(sSFR) between pairs and controls of this work (red) compared against those of \citet{Pan:2019} (Blue) and \citet{Thorp:2019} (Green).}
\label{fig:prof_comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In Figure \ref{fig:prof_comp} we look at the $\Delta$log(sSFR) profile as a function of galactocentric radius between our work and previous works. \citet{Pan:2019} examines a set of paired galaxies in the MaNGA survey while \citet{Thorp:2019} examines a set of post-merger galaxies in the MaNGA survey. The galaxy pairs in \citet{Pan:2019} have sSFR enhancements that are roughtly $\sim$0.1$-$0.2 dex higher than our sample. The $\Delta$log(sSFR) profile is closest to our work at the centers of the galaxy pairs but then begins to diverge further out into the galaxies' disks. Our paired galaxies show a slight sSFR suppression in their disks past 1.0 R/\reff\ while the galaxy pairs in \citet{Pan:2019} show an enhancement of $\sim$0.15 dex to the sSFR in their disks.

The post-merger galaxies in \citet{Thorp:2019} have sSFR enhancements between $\sim$0.05$-$0.2 higher than our sample of paired galaxies. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(sSFR) $=$ 0.5, roughly 0.25 dex higher than our paired galaxies. The post-merger galaxies also feature enhanced sSFR in their disk past 1.0 R/\reff\ of $\sim$0.1$-$0.15 dex. The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

\citet{Barrera-Ballesteros:2015} studied the sSFR in paired galaxies in the CALIFA IFS survey using the H$\alpha$ equivalent width (\ewha) as a proxy. They compared the sSFR in the centers of paired galaxies to their disks by varying the aperture diameters from which they extract the \ewha. They found that the sSFR is enhanced in the centers of the paired galaxies but suppressed in their disks. Between our work, \citet{Barrera-Ballesteros:2015}, and \citet{Pan:2019} there three is different scenarios; one where the disks of the paired galaxies see no change to the sSFR, one where the sSFR is suppressed, and one where the sSFR is enhanced.

Looking at individual difference profiles in our work, we see that there is a large variance in the $\Delta$log(sSFR) where there are as many paired galaxies with sSFR enhancement in their disks as there are paired galaxies with suppression in there disks. The end result is that the average of the $\Delta$log(sSFR) is zero. The merger induced effects seen in the disks of the paired galaxies may be more complicated than what is seen in the centers of the galaxies.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/prof_comp.pdf}
\caption[]{The mean radial profile of $\Delta$log(sSFR) between pairs and controls of this work (Black) compared against those of \citet{Pan:2019} (Red) and \citet{Thorp:2019} (Blue).}
\label{fig:prof_comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

In Figure \ref{fig:prof_comp} we compare the $\Delta$log(sSFR) profile as a function of galactocentric radius between our work and previous works. \citet{Pan:2019} examines a set of paired galaxies in the MaNGA survey while \citet{Thorp:2019} examines a set of post-merger galaxies in the MaNGA survey. The galaxy pairs in \citet{Pan:2019} have sSFR enhancements that are roughly $\sim$0.1$-$0.2 dex higher than our sample. Our paired galaxies show a slight sSFR suppression in their disks past 1.0 R/\reff\ while the galaxy pairs in \citet{Pan:2019} show an enhancement of $\sim$0.15 dex to the sSFR in their disks. The reason for the discrepancies between our samples is that \citet{Pan:2019} includes a set of post-merger galaxies while our sample has no post-merger galaxies. These post-merger galaxies are predicted to have higher levels of star formation over other paired galaxies due to a second larger burst of star formation which is triggered as the merging galaxies coalesce.

The post-merger galaxies in \citet{Thorp:2019} have sSFR enhancements between $\sim$0.05$-$0.2 dex higher than our sample of paired galaxies. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(sSFR) $=$ 0.5, roughly 0.25 dex higher than our paired galaxies. The post-merger galaxies also feature enhanced sSFR in their disk past 1.0 $R$/\reff\ of $\sim$0.1$-$0.15 dex. The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

\citet{Barrera-Ballesteros:2015} studied the sSFR in paired galaxies in the CALIFA IFS survey using the H$\alpha$ equivalent width (\ewha) as a proxy for the sSFR. They compared the sSFR in the centers of paired galaxies to their disks by varying the aperture diameters from which they extract the \ewha. They found that the sSFR is enhanced in the centers of the paired galaxies but suppressed in their disks, which is similar to what we find.
